---
title: "Homework 3"
author: "Shritama Ray"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Due date

Due: October 15 at 11:59pm. 

### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

### Problem 1: Instacart

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2: Accelerometer

#### Load and tody the data

```{r load accel}
accel_data = read_csv("./Data/accel_data.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "count") %>%
  mutate(day_type = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"))
```

#### Describe the dataset
This dataset has `r nrow(accel_data)` rows or observations and `r ncol(accel_data)` columns. The variables include the week number (week), a day id (day_id), the type of (day_type), day of the week (day), the minute of the day (minute), and the activity level recorded at each minute of the day (count).

#### Total activity per day

```{r table by day}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(day) %>%
  summarize(total_activity_count = sum(count, na.rm = TRUE)) %>%
  knitr::kable(caption = "*Table 1: Total Activity Count Per Day*")
```

*TRENDS:*
It seems like activity levels are highest during the middle/end of the week (Wednesday-Friday) with the lowest activity on Saturdays. 

We can check if this trend is consistent across all weeks:

```{r total per day per week}
accel_data %>%
  mutate(day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))) %>%
  group_by(week, day) %>%
  summarize(total_activity_count = sum(count, na.rm = TRUE)) %>%
  pivot_wider(names_from = day,values_from = total_activity_count
    ) %>%
  knitr::kable(caption = "*Table 2: Total Activity Count Per Day Per Week*")
```

*Comment:* It looks like the trend is NOT consistent across all weeks. For example, in Week 2, activity levels were highest on Saturday, while in Weeks 4 &5, activity levels were actually lowest on Saturday. There is quite a bit of variation across both weeks and days.  

#### 24-hour Activity
```{r 24 hour plot}
accel_data %>%
  mutate(minute = as.integer(minute))%>%
  group_by(day, minute) %>%
  ggplot(aes(x = minute, y = count, color = day, group = day_id)) + geom_point() + 
  labs(
    title = "24-hour Activity Per Day",
    x = "Hour(0-24)",
    y = "Activity") + 
  scale_x_continuous(
    breaks = c(0, 120, 240, 360, 480, 600, 720, 840, 960, 1080, 1200, 1320, 1440),
    labels = c("0", "2", "4", "6", "8", "10", "12", "14", "16", "18", "20", "22", "24"))
```


